workshop

### 2022-09-12 scripts/workshop.Rmd
really very confusing stuff

@@

```{r}
log_abun_mat <- log1p(abundance_mat)

@@

bc_dist_mat <- as.matrix(bc_dist)
bc_dist_df <- melt(bc_dist_mat, varnames=c("row", "col"))

ggplot(bc_dist_df, aes(x = col, y = row, fill = value)) + geom_tile()
```

>>>>

# Questions

Do communities that are similar in community composition have similar community competence values?

Can communities with different community compositions have similar community competence values?

How are these communities distributed across space and time?
i.e. how much spatial and temporal autocorrelation is there between sites in terms of community composition and community competence?

@@

Use Bray-Curtis distances to analyze how similar sites are to each other
Higher BC distance means more dissimilar
1 means perfect mismatch
```{r}
log_abun_mat <- log1p(abundance_mat)

@@

bc_dist_mat <- as.matrix(bc_dist)
bc_dist_df <- melt(bc_dist_mat, varnames=c("row", "col"), value.name = "bc_dist")

ggplot(bc_dist_df, aes(x = col, y = row, fill = bc_dist)) + geom_tile()
```

Manhattan method is the absolute value of the difference between two values
So higher distance means larger difference
Normalized to have range 0 to 1
```{r}
cc_diff <- comm_summ %>% dplyr::select(siteID, cc) %>% mutate(log_cc = log1p(cc)) %>% dplyr::select(-cc)
cc_dist <- dist(cc_diff, method = "manhattan")
cc_diff_mat <- as.matrix(cc_dist, labels = T)
colnames(cc_diff_mat) <- cc_diff[[1]]
rownames(cc_diff_mat) <- cc_diff[[1]]
cc_diff_df <- melt(cc_diff_mat, varnames=c("row", "col"), value.name = "cc_dist") %>% 
  mutate(cc_norm = (cc_dist-min(.$cc_dist))/(max(.$cc_dist)-min(.$cc_dist)))

ggplot(cc_diff_df, aes(x = col, y = row, fill = cc_norm)) + geom_tile()
```

```{r}
bc_cc_df <- left_join(cc_diff_df, bc_dist_df)
bc_cc_df %<>% mutate(diff = abs(bc_dist-cc_norm))

ggplot(bc_cc_df, aes(x = col, y = row, fill = diff)) + geom_tile()

bc_cc_df %>% filter(bc_dist!=1) %>% ggplot(.,aes(x=cc_norm, y=bc_dist)) + geom_point() + geom_smooth(method = "lm")
```

Principal Component Ranking may not be reliable as a proxy for community similarity because of double-zeroes

If two sites both have zeroes for a species then that does not show that a site is similar but only shows that both sites may be unsuitable for the species

Bray-Curtis dissimilarity does not face this limitation

But the question is how similar are sites with the similar community competence?

The manhattan distance is equivalent to bray-curtis distance
